Multidimensional Persistence Module Classification via Lattice-Theoretic Convolutions

### 2020-10-10 paper/paper.tex
added figures final edits/cuts

%HANS's CONVENIENCES
\usepackage{mymacros}

@@

Due to space constraints, we can offer only a brief overview of multiparameter
persistent homology. For a primer on persistent homology,
see~\cite{ghrist_barcodes:_2008, carlsson_topology_2009}; for multiparameter
persistent homology, see~\cite{lesnick_interactive_2015}. An introduction to
lattices may be found in~\cite{}.

@@

\subsection{Multiparameter persistence}
The Rips construction produces a filtration of simplicial complexes from a
finite metric space; it is natural to consider the behavior of the homology
functor over a pair of coherent filtrations. Consider
a finite metric space $(\mathcal M, d)$ and a filtration
function $\rho:\mathcal M \to \R$. This data specifies a bifiltration of
simplicial complexes given by
\[\mathbb{X}_{r,t} = \text{Rips}_r{x \in \mathcal M \mid \rho(x) \leq t}.\]
There is a natural inclusion $\mathbb{X}_{r,t} \hookrightarrow
\mathbb{X}_{r',t'}$ whenever $(r,t) \leq (r',t')$ in the lattice $\R \times R$.
Composing with the homology functor produces a 2-parameter persistence module
\[PH_{i}: \R_+ \times \R \to \Vect; \quad (r,t) \mapsto H_i(X_{r,t}).\]

@@

A lattice convolution layer takes as input an $N_{\text{in}}$-dimensional signal
$f : [m] \times [n] \to \R^{N_{\text{in}}}$ and outputs an
$N_{\text{out}}$-dimensional signal $ [m] \times [n] \to
\R^{N_{\text{out}}}$. The layer's parameters are given by a function $g: [m]
\times [n] \to \R^{N_{\text{out}}\times N_{\text{in}}}$. If we label the
entries of $f(x,y)$ by $f_i$ and the entries of $g(x,y)$ by $g^i_j$ The layer then acts by
\[\text{MeetConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\wedge} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \wedge a, y \wedge b)g^i_j(a,b)\]
in the case of convolution with respect to the meet operation, and
\[\text{JoinConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\vee} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \vee a, y \vee b)g^i_j(a,b)\]
in the case of convolution with respect to the join operation.

@@

\section{Experiments}
\begin{figure}
  \caption{Comparison training curves for the lattice neural network and
    standard convolutional neural network.}\label{fig:comparison}
\end{figure}

@@

The lattice-based convolution and classical convolution networks perform
similarly on the training set. However, the lattice convolution network's
predictions generalize more accurately to the testing set.
%% more discussion of what the data actually looks like

>>>>

%HANS's CONVENIENCES
\usepackage{mymacros}
\usepackage{graphicx}

@@

Due to space constraints, we can offer only a brief overview of multiparameter
persistent homology. For a primer on persistent homology,
see~\cite{ghrist_barcodes:_2008, carlsson_topology_2009}; for multiparameter
persistent homology, see~\cite{lesnick_interactive_2015}. An introduction to
lattices may be found in~\cite{davey2002introduction}.

@@

\subsection{Multiparameter persistence}
The Rips construction produces a filtration of simplicial complexes from a
finite metric space; it is natural to consider the behavior of the homology
functor over a pair of coherent filtrations. Consider
a finite metric space $(\mathcal M, d)$ and a filtration
function $\rho:\mathcal M \to \R$. This data specifies a bifiltration of
simplicial complexes given by
\[\mathbb{X}_{r,t} = \text{Rips}_r\{x \in \mathcal M \mid \rho(x) \leq t\}.\]
There is a natural inclusion $\mathbb{X}_{r,t} \hookrightarrow
\mathbb{X}_{r',t'}$ whenever $(r,t) \leq (r',t')$ in the lattice $\R \times R$.
Composing with the homology functor produces a 2-parameter persistence module
\[PH_{i}: \R_+ \times \R \to \Vect; \quad (r,t) \mapsto H_i(X_{r,t}).\]

@@

A lattice convolution layer takes as input an $N_{\text{in}}$-dimensional signal
$f : [m] \times [n] \to \R^{N_{\text{in}}}$ and outputs an
$N_{\text{out}}$-dimensional signal $ [m] \times [n] \to
\R^{N_{\text{out}}}$. The layer's parameters are given by a function $g: [m]
\times [n] \to \R^{N_{\text{out}}\times N_{\text{in}}}$. If we label the
entries of $f(x,y)$ by $f_i$ and the entries of $g(x,y)$ by $g^i_j$, the layer then acts by
\[\text{MeetConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\wedge} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \wedge a, y \wedge b)g^i_j(a,b)\]
in the case of convolution with respect to the meet operation, and
\[\text{JoinConv}(f)(x,y)^j = \sum_{i} (f_i \ast_{\vee} g^i_j)(x,y) = \sum_i
  \sum_{(a,b) \in [m]\times [n]} f_i(x \vee a, y \vee b)g^i_j(a,b)\]
in the case of convolution with respect to the join operation.

@@

\section{Experiments}
\begin{figure}[b]
  \caption{Comparison training curves for the lattice neural network and
    standard convolutional neural network.}\label{fig:comparison}
    \begin{center}
    \includegraphics[width=0.49\textwidth]{lattice_accuracy.png}
    \vspace{0.2cm}
    \includegraphics[width=0.49\textwidth]{conv_accuracy.png}
    \end{center}

\end{figure}

@@

The lattice-based convolution and classical convolution networks perform
similarly on the training set. However, the lattice convolution network's
predictions generalize more accurately to the testing set ($\approx 7\%$ improvement in testing accuracy).
This improvement is  at the expense of noisy training as well as dramatically increased training time.  
%% more discussion of what the data actually looks like
